---
title: "bm1 final project (new )"
author: "chianti shi"
date: "12/11/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(haven)
library(rvest)
library(stringr)
library(forcats)
library(knitr)
library(ggplot2)
library(ggpubr)
library(leaps)

```


```{r}
hospital_data = readxl::read_excel("./data/GHProject_Dataset.xlsx") %>%
  clean_names() %>% 
  mutate(patientid = as.character(patientid)) 

hospital_visit = hospital_data %>% 
  group_by(patientid) %>% 
  summarize(visit_num = n()) %>% 
  inner_join(hospital_data, hospital_visit, by = "patientid") %>% 
  separate(., admitdtm, into = c("dayofweek","date","year"), sep = ",") %>% 
  mutate(visit_num = as.numeric(visit_num)) 

remove_id = hospital_visit %>% 
 mutate(date = str_replace(date, "January ", "01"),
        date = str_replace(date, "February ", "02"),
        date = str_replace(date, "March ", "03"),
        date = str_replace(date, "April ", "04"),
        date = str_replace(date, "May ", "05"),
        date = str_replace(date, "June ", "06"),
        date = str_replace(date, "July ", "07"),
        date = str_replace(date, "August ", "08"),
        date = str_replace(date, "September ", "09"),
        date = str_replace(date, "October ", "10"),
        date = str_replace(date, "November ", "11"),
        date = str_replace(date, "December ", "12")) %>% 
  mutate(date = as.numeric(date)) %>% 
  arrange(date) %>% 
  group_by(patientid) %>% 
  filter(date == min(date)) %>% 
  dplyr::select(-visit_num, -dayofweek, -facilityname, -postalcode, -facilityzip, -mews) %>% 
  mutate(visitid = as.character(visitid),
         is30dayreadmit = as.character(is30dayreadmit),
        cindex = as.character(cindex),
        icu_flag = as.character(icu_flag))

```

```{r}

summary(remove_id)
```


```{r}
par(mfrow=c(2,3))
hist(remove_id$loshours)
hist(remove_id$losdays2)
hist(remove_id$ageyear)
hist(remove_id$bmi)
hist(remove_id$bpsystolic)
hist(remove_id$o2sat)
hist(remove_id$temperature)
hist(remove_id$heartrate)
hist(remove_id$respirationrate)
hist(remove_id$bpdiastolic)
```

We observe that `loshours`, `losdays2`, `evisit` lack normality so we try to do (I tried a couple transformations and was unable to make evisit normal)


```{r warning = FALSE}
log_loshours = ggplot(data = remove_id, aes(x = log(loshours))) +
  geom_histogram() 

log_losdays2 = ggplot(data = remove_id, aes(x = log(losdays2))) +
  geom_histogram()

ggarrange(log_loshours, log_losdays2 + rremove("x.text"), 
          labels = c("A", "B"),
          ncol = 2, nrow = 1)
```

```{r}
patient_data = remove_id %>% 
  mutate(loshours = log(loshours),
         losdays2 = log(losdays2)) %>% 
  mutate(evisit, evisit = ifelse( evisit>0, 1, evisit)) %>% 
  mutate(evisit = as.character(evisit))

replace_missing = function(vector, func) {
  
  if(is.numeric(vector)) {
   func_vec = func(vector, na.rm = TRUE)
   vector[is.na(vector)] = func_vec
  }
  return(vector)
}

patient_data1 = map_df(patient_data, ~replace_missing(.x, func = mean)) %>% 
  na.omit() 
```


There are 77 in `maritalstatus`, 34 in `insurancetype`, 685 in `bmi`, 5 in `bpsystolic`, 3 in `o2sat`, 2 in `temperature`, 5 in `heartrate`, 3 in `respirationrate`, 1 in `bpdiastolic`. For missing values in numeric varaibles, NA are replaced by the mean of the rest of the dataset, NA in character variables are removed. 


###catagorical: 

```{r}
patient_data1 %>% 
  select(is30dayreadmit, cindex, evisit, icu_flag, gender, race, religion, maritalstatus, insurancetype) %>% 
  table()
```


