---
title: "bm1 final project (new )"
author: "chianti shi"
date: "12/11/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(haven)
library(rvest)
library(stringr)
library(forcats)
library(knitr)
library(ggplot2)
library(ggpubr)
library(leaps)

```


```{r}
hospital_data = readxl::read_excel("./data/GHProject_Dataset.xlsx") %>%
  clean_names() %>% 
  mutate(patientid = as.character(patientid)) 

hospital_visit = hospital_data %>% 
  group_by(patientid) %>% 
  summarize(visit_num = n()) %>% 
  inner_join(hospital_data, hospital_visit, by = "patientid") %>% 
  separate(., admitdtm, into = c("dayofweek","date","year"), sep = ",") %>% 
  mutate(visit_num = as.numeric(visit_num)) 

remove_id = hospital_visit %>% 
 mutate(date = str_replace(date, "January ", "01"),
        date = str_replace(date, "February ", "02"),
        date = str_replace(date, "March ", "03"),
        date = str_replace(date, "April ", "04"),
        date = str_replace(date, "May ", "05"),
        date = str_replace(date, "June ", "06"),
        date = str_replace(date, "July ", "07"),
        date = str_replace(date, "August ", "08"),
        date = str_replace(date, "September ", "09"),
        date = str_replace(date, "October ", "10"),
        date = str_replace(date, "November ", "11"),
        date = str_replace(date, "December ", "12")) %>% 
  mutate(date = as.numeric(date)) %>% 
  arrange(date) %>% 
  group_by(patientid) %>% 
  filter(date == min(date)) %>% 
  dplyr::select(-visit_num, -dayofweek, -facilityname, -postalcode, -facilityzip, -mews) %>% 
  mutate(visitid = as.character(visitid),
         is30dayreadmit = as.character(is30dayreadmit),
        cindex = as.character(cindex),
        icu_flag = as.character(icu_flag))

```

```{r}

summary(remove_id)
```


```{r}
par(mfrow=c(2,3))
hist(remove_id$loshours)
hist(remove_id$losdays2)
hist(remove_id$ageyear)
hist(remove_id$bmi)
hist(remove_id$bpsystolic)
hist(remove_id$o2sat)
hist(remove_id$temperature)
hist(remove_id$heartrate)
hist(remove_id$respirationrate)
hist(remove_id$bpdiastolic)
```

We observe that `loshours`, `losdays2`, `evisit` lack normality so we try to do (I tried a couple transformations and was unable to make evisit normal)


```{r warning = FALSE}
log_loshours = ggplot(data = remove_id, aes(x = log(loshours))) +
  geom_histogram() 

log_losdays2 = ggplot(data = remove_id, aes(x = log(losdays2))) +
  geom_histogram()

ggarrange(log_loshours, log_losdays2 + rremove("x.text"), 
          labels = c("A", "B"),
          ncol = 2, nrow = 1)
```

```{r}
patient_data = remove_id %>% 
  mutate(loshours = log(loshours),
         losdays2 = log(losdays2)) %>% 
  mutate(evisit, evisit = ifelse( evisit>0, 1, evisit)) %>% 
  mutate(evisit = as.character(evisit))

replace_missing = function(vector, func) {
  
  if(is.numeric(vector)) {
   func_vec = func(vector, na.rm = TRUE)
   vector[is.na(vector)] = func_vec
  }
  return(vector)
}

patient_data1 = map_df(patient_data, ~replace_missing(.x, func = mean)) %>% 
  na.omit() 
```


There are 77 in `maritalstatus`, 34 in `insurancetype`, 685 in `bmi`, 5 in `bpsystolic`, 3 in `o2sat`, 2 in `temperature`, 5 in `heartrate`, 3 in `respirationrate`, 1 in `bpdiastolic`. For missing values in numeric varaibles, NA are replaced by the mean of the rest of the dataset, NA in character variables are removed. 


###catagorical: 

```{r}
patient_data1 = patient_data1 %>% 
  mutate(cindex = recode(cindex, '0' = "Normal", '1' = "Mild", '2' = "Mild", '3' = "Moderate and Severe", '5' = "Moderate and Severe"),
         race = recode(race, "White" = "White", "African Amer/Black" = "African Amer/Black", "Asian" = "Other", "Native Amer/Alaskan" = "Other", "Natv Hawaii/Pacf Isl" = "Other", "Other/Multiracial" = "Other"),
         religion = recode(religion, "Catholic" = "Catholic", "Christian" = "Christian", "Jewish" = "Other", "No Affiliation" = "Other", "Other" = "Other", "Islam" = "Other", "Hindu" = "Other", "Mormon" = "Other", "Angelican" = "Other", "Hebrew" = "Other", "Non Denominational" = "Other"),
         maritalstatus = recode(maritalstatus, "Single" = "Single", "Married" = "Married", "Widowed" = "Other", "Divorced" = "Other", "Separated" = "Other", "Civil Union" = "Other"))


lm_cindex = lm(losdays2 ~ cindex, data = patient_data1)
lm_race = lm(losdays2 ~ race, data = patient_data1)
lm_religion = lm(losdays2 ~ religion, data = patient_data1)
lm_maritalstatus = lm(losdays2 ~ maritalstatus, data = patient_data1)

lm_evisit = lm(losdays2 ~ evisit, data = patient_data1)
lm_icu_flag = lm(losdays2 ~ icu_flag, data = patient_data1)
lm_gender = lm(losdays2 ~ gender, data = patient_data1)
lm_is30dayreadmit = lm(losdays2 ~ is30dayreadmit, data = patient_data1)
lm_insurancetype = lm(losdays2 ~ insurancetype, data = patient_data1)

summary(lm_cindex)
summary(lm_race)
summary(lm_religion)
summary(lm_maritalstatus)

summary(lm_evisit)
summary(lm_icu_flag)
summary(lm_gender)
summary(lm_is30dayreadmit)
summary(lm_insurancetype)


patient_data1 = patient_data1 %>% 
  select(-c(race, gender, religion)) %>% 
  filter(temperature > 25 & temperature< 45) %>%  #https://en.wikipedia.org/wiki/Human_body_temperature According to Wikkipage, the maxium and minimum temperature which human can endure.
  filter(heartrate<220)
```

```{r include=FALSE}
#Numerical:  ageyear, bmi, "bpsystolic"  ,  "o2sat"  ,temperature" , "heartrate"       "respirationrate", "bpdiastolic"
# data analysis
patient_data1 %>% 
  select(ageyear, bmi, bpsystolic, o2sat, temperature, heartrate, respirationrate, bpdiastolic) %>%  
  gather(key = "variables", value = "number", ageyear:bpdiastolic) %>% 
  group_by(variables) %>% 
  summarise(n_obs = n(),
             mean = round(mean(number), digits = 2),
               sd = round(sd(number), digits = 2),
           median = median(number),
          munimum = min(number),
           maxiun = max(number),
        first_Qu. = quantile(number, 0.25),
        third_Qu. = quantile(number, 0.75)
           )
```

```{r}
model = lm(losdays2 ~ ageyear+ bmi+ bpsystolic+ o2sat + temperature + heartrate + respirationrate +bpdiastolic + cindex + evisit + icu_flag + maritalstatus + insurancetype + is30dayreadmit, data = patient_data1 )

best <- function(model, ...) 
{
  subsets <- regsubsets(formula(model), model.frame(model), ...)
  subsets <- with(summary(subsets),
                  cbind(p = as.numeric(rownames(which)), which, rss, rsq, adjr2, cp, bic))
  
  return(subsets)
} 

round(best(model, nvmax= 30, nbest = 1), 4)

```


**1.Forward elimination procedures**
```{r}
model = lm(losdays2 ~ ageyear+ bmi+ bpsystolic+ o2sat + temperature + heartrate + respirationrate +bpdiastolic + cindex + evisit + icu_flag + maritalstatus + insurancetype + is30dayreadmit, data = patient_data1 )
#ols_step_forward(model, details = TRUE)

# result: ageyear is30dayreadmit gender insurancetype race bmi
```


**2.Backward elimination procedures**
```{r}
model = lm(losdays2 ~ ageyear+ bmi+ bpsystolic+ o2sat + temperature + heartrate + respirationrate +bpdiastolic + cindex + evisit + icu_flag + maritalstatus + insurancetype + is30dayreadmit, data = patient_data1 )
#ols_best_subset(model)

```

**3.Stepwise regression procedures**
```{r}
model = lm(losdays2 ~ ageyear+ bmi+ bpsystolic+ o2sat + temperature + heartrate + respirationrate +bpdiastolic + cindex + evisit + icu_flag + maritalstatus + insurancetype + is30dayreadmit, data = patient_data1 )
#ols_stepwise(model, details = FALSE)
##not work
```






###Anova

```{r}
mlr_automatic = lm(losdays2 ~ is30dayreadmit+cindex+evisit+ageyear+maritalstatus+insurancetype+bpsystolic+temperature+heartrate+respirationrate+bpdiastolic, data = patient_data1)  


mlr_criterion = lm(losdays2 ~ is30dayreadmit+cindex+evisit+ageyear+maritalstatus+insurancetype+bpsystolic+temperature+heartrate+respirationrate+bpdiastolic+o2sat, data = patient_data1)  
```

```{r}
summary(mlr_automatic)
summary(mlr_criterion)
```


```{r}
anova(mlr_automatic, mlr_criterion)
```
we choose the results with `mlr_automatic`, since the p value of 0.1179 indicating that  `mlr_criterion` is not superior. Also when we look at the summary for  `mlr_criterion` , the variable `o2sat` is not significant. 


```{r}
par(mfrow=c(2,2))
plot(mlr_automatic)
par(mfrow=c(2,2))
plot(mlr_criterion)
```

(a) Measures of influence : identify any leverage

We first check outliers in Y:

```{r}
influence_measure = rstandard(mlr_automatic)
outlier_y = influence_measure[abs(influence_measure) > 2.5]
outlier_y
```

We then get hat matrix to check outliers in X:
```{r}
hat = hatvalues(mlr_automatic)
hat[hat > 0.2]
hat[hat > 2*2/4]
```

```{r}
nooutliers <- patient_data1[-c(25,35,118,152,157,220,223,226,445,498,570,593,814,839,841,872,915,916,1029,1124,1134,1248,1255,1256,1341,1348,1473,1494,1677,1799,1811,1838,2051,2052,2117,2233,2251,2261,2485,2489,2504,2582,2605,2616,2717,2752,2796,2797,2831,2858,2885,2929,2970,3177,3199,3265,3292),]

mult.fit_nooutliers<- lm(losdays2 ~ is30dayreadmit+cindex+evisit+ageyear+maritalstatus+insurancetype+bpsystolic+temperature+heartrate+respirationrate+bpdiastolic, data = nooutliers) 
summary(mult.fit_nooutliers)
summary(mult.fit_nooutliers)
```


```{r}
set.seed(1)

  patient_data1 %>% 
  select(losdays2, is30dayreadmit, cindex, evisit, ageyear, maritalstatus, insurancetype, bpsystolic, temperature, heartrate, respirationrate, bpdiastolic) %>% 
  modelr::bootstrap(n = 1000) %>% 
  mutate(models = map(strap, ~lm(losdays2 ~ is30dayreadmit+cindex+evisit+ageyear+maritalstatus+insurancetype+bpsystolic+temperature+heartrate+respirationrate+bpdiastolic, data = .x)),
         results = map(models, broom::tidy)) %>% 
  select(-strap, -models) %>% 
  unnest() %>%
  group_by(term) %>% 
  summarize(mean_estimate = round(mean(estimate), 2),
            variance = round(var(estimate), 2))
  
  # compare to model estimates
  summary(mlr_automatic)
```


