---
title: "p8130 final project. Muhire's Branch"
author: "chianti shi"
date: "12/8/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(haven)
library(rvest)
library(stringr)
library(forcats)
library(knitr)
library(ggplot2)
library(ggpubr)
library(dplyr)

```


Loading and cleaning the dataset
```{r, include=FALSE}
hospital_data = readxl::read_excel("./data/GHProject_Dataset.xlsx") %>%
  clean_names() %>% 
  mutate(patientid = as.character(patientid)) 

hospital_visit = hospital_data %>% 
  group_by(patientid) %>% 
  summarize(visit_num = n()) %>% 
  inner_join(hospital_data, hospital_visit, by = "patientid") %>% 
  separate(., admitdtm, into = c("dayofweek","date","year"), sep = ",") %>% 
  mutate(visit_num = as.numeric(visit_num)) 

remove_id = hospital_visit %>% 
 mutate(date = str_replace(date, "January ", "01"),
        date = str_replace(date, "February ", "02"),
        date = str_replace(date, "March ", "03"),
        date = str_replace(date, "April ", "04"),
        date = str_replace(date, "May ", "05"),
        date = str_replace(date, "June ", "06"),
        date = str_replace(date, "July ", "07"),
        date = str_replace(date, "August ", "08"),
        date = str_replace(date, "September ", "09"),
        date = str_replace(date, "October ", "10"),
        date = str_replace(date, "November ", "11"),
        date = str_replace(date, "December ", "12")) %>% 
  mutate(date = as.numeric(date)) %>% 
  arrange(date) %>% 
  group_by(patientid) %>% 
  filter(date == min(date)) 
```

Select variables of interest:
```{r}
visit_data = remove_id %>% 
  dplyr::select(-visit_num, -dayofweek, -facilityname, -postalcode, -facilityzip) %>% 
  mutate(is30dayreadmit = as.character(is30dayreadmit),
         cindex = as.character(cindex),
         icu_flag = as.character(icu_flag),
         mews = as.character(mews))
  
```


We do not check normality for variables `gender`, `race`, `religion`, `maritalstatus`,`insurancetype` as they are catagorical variables. 

subset for varaible summary: 
```{r}
summary_data = visit_data %>% 
  dplyr::select(-gender, -race, -religion, -maritalstatus, -insurancetype,  -visitid, -date, -year, -is30dayreadmit, -cindex, -icu_flag, -mews)
```

```{r}
summary(summary_data)
```


```{r}
par(mfrow = c(2, 3))
hist(visit_data$loshours)
hist(visit_data$losdays2)
hist(visit_data$evisit)
hist(visit_data$ageyear)
hist(visit_data$bmi)
hist(visit_data$bpsystolic)
hist(visit_data$o2sat)
hist(visit_data$temperature)
hist(visit_data$heartrate)
hist(visit_data$respirationrate)
hist(visit_data$bpdiastolic)
```


We observe that `loshours`, `losdays2`, `evisit` lack normality so we try to do (I tried a couple transformations and was unable to make evisit normal)

```{r warning = FALSE}
log_loshours = ggplot(data = visit_data, aes(x = log(loshours))) +
  geom_histogram() 

log_losdays2 = ggplot(data = visit_data, aes(x = log(losdays2))) +
  geom_histogram()

ggarrange(log_loshours, log_losdays2 + rremove("x.text"), 
          labels = c("A", "B"),
          ncol = 2, nrow = 1)
```


```{r}
patient_data = visit_data %>%
  mutate(loshours = log(loshours),
  losdays2 = log(losdays2))
  
  replace_missing = function(vector, func) {
  if (is.numeric(vector)) {
  func_vec = func(vector, na.rm = TRUE)
  vector[is.na(vector)] = func_vec
  }
  return(vector)
  }
  
  patient_data1 = map_df(patient_data, ~ replace_missing(.x, func = mean)) %>%
  na.omit()

```
There are 163 NA in `mews`, 77 in `maritalstatus`, 34 in `insurancetype`, 685 in `bmi`, 5 in `bpsystolic`, 3 in `o2sat`, 2 in `temperature`, 5 in `heartrate`, 3 in `respirationrate`, 1 in `bpdiastolic`. For missing values in numeric varaibles, NA are replaced by the mean of the rest of the dataset, NA in character variables are removed. 


you guys can use this `patient_data1` to perform further tests except the issue with untransformed `evisit`. 

```{r}
# Saving dataset for automatic search procedures in SAS

write_csv(patient_data1, "/Users/mhk/Desktop/patient_data1.csv")

# patient_data1 = patient_data1 %>%
#   mutate(male = ifelse(gender == "Male", 1, 0),
#          afr_american = ifelse(race == "African Amer/Black", 1, 0),
#          asian = ifelse(race == "Asian", 1, 0),
#          nat_american = ifelse(race == "Native Amer/Alaskan", 1, 0),
#          nat_hawaii = ifelse(race == "Natv Hawaii/Pacf Isl", 1, 0),
#          multi_racial = ifelse(race == "Other/Multiracial", 1, 0),
#          angelican = ifelse(religion == "Angelican", 1, 0),
#          catholic = ifelse(religion == "Catholic", 1, 0),
#          christian = ifelse(religion == "Christian", 1, 0),
#          hebrew = ifelse(religion == "Hebrew", 1, 0),
#          hindu = ifelse(religion == "Hindu", 1, 0),
#          islam = ifelse(religion == "Islam", 1, 0),
#          jewish = ifelse(religion == "Jewish", 1, 0),
#          mormon = ifelse(religion == "Mormon", 1, 0),
#          non_denom = ifelse(religion == "Non Denominational", 1, 0),
#          married = ifelse(religion == "Married", 1, 0),
#          widowed = ifelse(maritalstatus == "Widowed", 1, 0),
#          divorced = ifelse(religion == "Divorced", 1, 0),
#          separated = ifelse(religion == "Separated", 1, 0),
#          
#          )
  
  # mutate(gender = recode(gender, "Female" = 1, "Male" = 0),
  #        race = recode(race, "White" = 0, "African Amer/Black" = 1, "Asian" = 2, "Native Amer/Alaskan" = 3, "Natv Hawaii/Pacf Isl" = 4, "Other/Multiracial" = 5),
  #        religion = recode(religion, "No Affiliation" = 0, "Angelican" = 1,  "Catholic" = 2, "Christian" = 3, "Hebrew" = 4, "Hindu" = 5, "Islam" = 6, "Jewish" = 7, "Mormon" = 8, "Non Denominational" = 9, "Other" = 10),
  #        maritalstatus = recode(maritalstatus, "Single" = 0, "Married" = 1, "Widowed" = 2, "Divorced" = 3, "Separated" = 4, "Civil Union" = 5),
  #       insurancetype = recode(insurancetype, "Private" = 0, "Medicare" = 1, "Medicaid" = 2)) %>%
  # write_csv(., "/Users/mhk/Desktop/patient_data1.csv")

```

