---
title: "p8130 final project"
author: "chianti shi"
date: "12/8/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(haven)
library(rvest)
library(stringr)
library(forcats)
library(knitr)
library(ggplot2)
library(ggpubr)
library(leaps)
```


Loading and cleaning the dataset
```{r, include=FALSE}
hospital_data = readxl::read_excel("./data/GHProject_Dataset.xlsx") %>%
  clean_names() %>% 
  mutate(patientid = as.character(patientid)) 

hospital_visit = hospital_data %>% 
  group_by(patientid) %>% 
  summarize(visit_num = n()) %>% 
  inner_join(hospital_data, hospital_visit, by = "patientid") %>% 
  separate(., admitdtm, into = c("dayofweek","date","year"), sep = ",") %>% 
  mutate(visit_num = as.numeric(visit_num)) 

remove_id = hospital_visit %>% 
 mutate(date = str_replace(date, "January ", "01"),
        date = str_replace(date, "February ", "02"),
        date = str_replace(date, "March ", "03"),
        date = str_replace(date, "April ", "04"),
        date = str_replace(date, "May ", "05"),
        date = str_replace(date, "June ", "06"),
        date = str_replace(date, "July ", "07"),
        date = str_replace(date, "August ", "08"),
        date = str_replace(date, "September ", "09"),
        date = str_replace(date, "October ", "10"),
        date = str_replace(date, "November ", "11"),
        date = str_replace(date, "December ", "12")) %>% 
  mutate(date = as.numeric(date)) %>% 
  arrange(date) %>% 
  group_by(patientid) %>% 
  filter(date == min(date)) 
```

Select variables of interest:
```{r}
visit_data = remove_id %>% 
  dplyr::select(-visit_num, -dayofweek, -facilityname, -postalcode, -facilityzip) %>% 
  mutate(is30dayreadmit = as.character(is30dayreadmit),
         cindex = as.character(cindex),
         icu_flag = as.character(icu_flag),
         mews = as.character(mews))
  
```


We do not check normality for variables `gender`, `race`, `religion`, `maritalstatus`,`insurancetype` as they are catagorical variables. 

subset for varaible summary: 
```{r}
summary_data = visit_data %>% 
  dplyr::select(-gender, -race, -religion, -maritalstatus, -insurancetype,  -visitid, -date, -year, -is30dayreadmit, -cindex, -icu_flag, -mews)
```

```{r}
summary(summary_data)
```


```{r}
par(mfrow=c(2,3))
hist(visit_data$loshours)
hist(visit_data$losdays2)
hist(visit_data$evisit)
hist(visit_data$ageyear)
hist(visit_data$bmi)
hist(visit_data$bpsystolic)
hist(visit_data$o2sat)
hist(visit_data$temperature)
hist(visit_data$heartrate)
hist(visit_data$respirationrate)
hist(visit_data$bpdiastolic)
```


We observe that `loshours`, `losdays2`, `evisit` lack normality so we try to do (I tried a couple transformations and was unable to make evisit normal)

```{r warning = FALSE}
log_loshours = ggplot(data = visit_data, aes(x = log(loshours))) +
  geom_histogram() 

log_losdays2 = ggplot(data = visit_data, aes(x = log(losdays2))) +
  geom_histogram()

ggarrange(log_loshours, log_losdays2 + rremove("x.text"), 
          labels = c("A", "B"),
          ncol = 2, nrow = 1)
```


```{r}
patient_data = visit_data %>% 
  mutate(loshours = log(loshours),
         losdays2 = log(losdays2)) %>% 
  mutate(mews = as.numeric(mews)) %>% 
#mutate mews column:   0, 1 to "1", 2,3,to "2", 4,5 to "3", 6 to "4"
# note: there are NAs in variable mews.
mutate(mews, mews = ifelse( mews>5, 6, mews)) %>% 
  mutate(mews = as.character(mews)) %>% 
  mutate(mews = recode(mews, `0` = "normal", `1` = "normal", `2` = "increase_caution", `3` = "increase_caution", `4` = "further_deterioration", `5` = "further_deterioration", `6` = "immediate_action_required")) %>% 
  mutate(mews = recode(mews, `normal` = "1", `increase_caution` = "2", `further_deterioration` = "3", `immediate_action_required` = "4" )) %>% 
 mutate(evisit, evisit = ifelse( evisit>0, 1, evisit)) %>% 
  mutate(evisit = as.character(evisit))

visit_data = remove_id %>% 
  dplyr::select(-visit_num, -dayofweek, -facilityname, -postalcode, -facilityzip) %>%
  mutate(mews, mews = ifelse( mews>5, 6, mews)) %>% 
  mutate(mews = as.character(mews)) %>% 
  mutate(mews = recode(mews, `0` = "normal", `1` = "normal", `2` = "increase_caution", `3` = "increase_caution", `4` = "further_deterioration", `5` = "further_deterioration", `6` = "immediate_action_required")) %>% 
  mutate(mews = recode(mews, `normal` = "1", `increase_caution` = "2", `further_deterioration` = "3", `immediate_action_required` = "4" )) %>% 
  mutate(mews = as.numeric(mews))

replace_missing = function(vector, func) {
  
  if(is.numeric(vector)) {
   func_vec = func(vector, na.rm = TRUE)
   vector[is.na(vector)] = func_vec
  }
  return(vector)
}

patient_data1 = map_df(patient_data, ~replace_missing(.x, func = mean)) %>% 
  na.omit() 

```
There are 163 NA in `mews`, 77 in `maritalstatus`, 34 in `insurancetype`, 685 in `bmi`, 5 in `bpsystolic`, 3 in `o2sat`, 2 in `temperature`, 5 in `heartrate`, 3 in `respirationrate`, 1 in `bpdiastolic`. For missing values in numeric varaibles, NA are replaced by the mean of the rest of the dataset, NA in character variables are removed. 


you guys can use this `patient_data1` to perform further tests except the issue with untransformed `evisit`. 



```{r}
##gender race religion maritalstatus
patient_data_lm = lm(losdays2 ~ is30dayreadmit + mews + cindex + evisit + icu_flag + ageyear + gender + race + religion + maritalstatus + insurancetype, data = patient_data1)

best <- function(model, ...) 
{
  subsets <- regsubsets(formula(model), model.frame(model), ...)
  subsets <- with(summary(subsets),
                  cbind(p = as.numeric(rownames(which)), which, rss, rsq, adjr2, cp, bic))
  
  return(subsets)
}  

round(best(patient_data_lm, nbest = 1), 4)    
```

