---
title: "p8130 final project"
author: "chianti shi"
date: "12/8/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(haven)
library(rvest)
library(stringr)
library(forcats)
library(knitr)
library(ggplot2)
library(ggpubr)
library(leaps)
```


Loading and cleaning the dataset
```{r, include=FALSE}
hospital_data = readxl::read_excel("./data/GHProject_Dataset.xlsx") %>%
  clean_names() %>% 
  mutate(patientid = as.character(patientid)) 

hospital_visit = hospital_data %>% 
  group_by(patientid) %>% 
  summarize(visit_num = n()) %>% 
  inner_join(hospital_data, hospital_visit, by = "patientid") %>% 
  separate(., admitdtm, into = c("dayofweek","date","year"), sep = ",") %>% 
  mutate(visit_num = as.numeric(visit_num)) 

remove_id = hospital_visit %>% 
 mutate(date = str_replace(date, "January ", "01"),
        date = str_replace(date, "February ", "02"),
        date = str_replace(date, "March ", "03"),
        date = str_replace(date, "April ", "04"),
        date = str_replace(date, "May ", "05"),
        date = str_replace(date, "June ", "06"),
        date = str_replace(date, "July ", "07"),
        date = str_replace(date, "August ", "08"),
        date = str_replace(date, "September ", "09"),
        date = str_replace(date, "October ", "10"),
        date = str_replace(date, "November ", "11"),
        date = str_replace(date, "December ", "12")) %>% 
  mutate(date = as.numeric(date)) %>% 
  arrange(date) %>% 
  group_by(patientid) %>% 
  filter(date == min(date)) 
```

Select variables of interest:
```{r}
visit_data = remove_id %>% 
  dplyr::select(-visit_num, -dayofweek, -facilityname, -postalcode, -facilityzip) %>% 
  mutate(is30dayreadmit = as.character(is30dayreadmit),
         cindex = as.character(cindex),
         icu_flag = as.character(icu_flag),
         mews = as.character(mews))
  
```


We do not check normality for variables `gender`, `race`, `religion`, `maritalstatus`,`insurancetype` as they are catagorical variables. 

subset for varaible summary: 
```{r}
summary_data = visit_data %>% 
  dplyr::select(-gender, -race, -religion, -maritalstatus, -insurancetype,  -visitid, -date, -year, -is30dayreadmit, -cindex, -icu_flag, -mews)
```

```{r}
summary(summary_data)
```


```{r}
par(mfrow=c(2,3))
hist(visit_data$loshours)
hist(visit_data$losdays2)
hist(visit_data$evisit)
hist(visit_data$ageyear)
hist(visit_data$bmi)
hist(visit_data$bpsystolic)
hist(visit_data$o2sat)
hist(visit_data$temperature)
hist(visit_data$heartrate)
hist(visit_data$respirationrate)
hist(visit_data$bpdiastolic)
```


We observe that `loshours`, `losdays2`, `evisit` lack normality so we try to do (I tried a couple transformations and was unable to make evisit normal)

```{r warning = FALSE}
log_loshours = ggplot(data = visit_data, aes(x = log(loshours))) +
  geom_histogram() 

log_losdays2 = ggplot(data = visit_data, aes(x = log(losdays2))) +
  geom_histogram()

ggarrange(log_loshours, log_losdays2 + rremove("x.text"), 
          labels = c("A", "B"),
          ncol = 2, nrow = 1)
```


```{r}
patient_data = visit_data %>% 
  mutate(loshours = log(loshours),
         losdays2 = log(losdays2)) %>% 
  mutate(mews = as.numeric(mews)) %>% 
#mutate mews column:   0, 1 to "1", 2,3,to "2", 4,5 to "3", 6 to "4"
# note: there are NAs in variable mews.
mutate(mews, mews = ifelse( mews>5, 6, mews)) %>% 
  mutate(mews = as.character(mews)) %>% 
  mutate(mews = recode(mews, `0` = "normal", `1` = "normal", `2` = "increase_caution", `3` = "increase_caution", `4` = "further_deterioration", `5` = "further_deterioration", `6` = "immediate_action_required")) %>% 
  mutate(mews = recode(mews, `normal` = "1", `increase_caution` = "2", `further_deterioration` = "3", `immediate_action_required` = "4" )) %>% 
 mutate(evisit, evisit = ifelse( evisit>0, 1, evisit)) 

patient_data = remove_id %>% 
  dplyr::select(-visit_num, -dayofweek, -facilityname, -postalcode, -facilityzip) %>%
  mutate(mews, mews = ifelse( mews>5, 6, mews)) %>% 
  mutate(mews = as.character(mews)) %>% 
  mutate(mews = recode(mews, `0` = "normal", `1` = "normal", `2` = "increase_caution", `3` = "increase_caution", `4` = "further_deterioration", `5` = "further_deterioration", `6` = "immediate_action_required")) %>% 
  mutate(mews = recode(mews, `normal` = "1", `increase_caution` = "2", `further_deterioration` = "3", `immediate_action_required` = "4" )) %>% 
  mutate(is30dayreadmit = as.character(is30dayreadmit),
         cindex = as.character(cindex),
         icu_flag = as.character(icu_flag),
         mews = as.character(mews),
         evisit = as.character(evisit))

replace_missing = function(vector, func) {
  
  if(is.numeric(vector)) {
   func_vec = func(vector, na.rm = TRUE)
   vector[is.na(vector)] = func_vec
  }
  return(vector)
}

patient_data1 = map_df(patient_data, ~replace_missing(.x, func = mean)) %>% 
  na.omit() 

```
There are 163 NA in `mews`, 77 in `maritalstatus`, 34 in `insurancetype`, 685 in `bmi`, 5 in `bpsystolic`, 3 in `o2sat`, 2 in `temperature`, 5 in `heartrate`, 3 in `respirationrate`, 1 in `bpdiastolic`. For missing values in numeric varaibles, NA are replaced by the mean of the rest of the dataset, NA in character variables are removed. 
you guys can use this `patient_data1` to perform further tests except the issue with untransformed `evisit`. 


```{r}
patient_data_lm = lm(losdays2 ~ is30dayreadmit + mews + cindex + bmi + evisit + icu_flag + ageyear + gender + race + religion + maritalstatus + insurancetype, data = patient_data1)
#gendermale  ageyear evisit234 cindex35 is30days

```

```{r}
attach(patient_data1)
```

```{r admission}
within30_lm = lm(patient_data1$losdays2~patient_data1$is30dayreadmit)
summary(within30_lm)
par(mfrow=c(2,2))
plot(within30)
```


```{r cindex}

cindex_lm = lm(patient_data1$losdays2~cindex)
summary(cindex_lm)
par(mfrow=c(2,2))
plot(cindex_lm)
```

Severity: 0=normal, 1-2=mild, 3-4=moderate and >5=severe 

```{r evisit}
 #don't know how to transform "evisit"


evisit_lm = lm(patient_data1$losdays2~evisit)
summary(evisit_lm)

```

```{r icu_flag}


icu_flag_lm = lm(patient_data1$losdays2~icu_flag)
summary(icu_flag_lm)

```

```{r age}
age_lm = lm(patient_data1$losdays2~ageyear)
summary(age_lm)
par(mfrow=c(2,2))
plot(age_lm)
```

```{r gender}
gender_lm = lm(patient_data1$losdays2~gender)
summary(gender_lm)
anova(gender_lm)
par(mfrow=c(2,2))
plot(gender_lm)
```
not significant

```{r race}
race_lm = lm(patient_data1$losdays2~race)
summary(race_lm)
anova(race_lm)
par(mfrow=c(2,2))
plot(race_lm)
```
not significant

```{r religion}
religion_lm = lm(patient_data1$losdays2~religion)
summary(religion_lm)
anova(religion_lm)
par(mfrow=c(2,2))
plot(religion_lm)
```
Not significant

```{r maritalstatus}
maritalstatus_lm = lm(patient_data1$losdays2~maritalstatus)
summary(maritalstatus_lm)
anova(maritalstatus_lm)
par(mfrow=c(2,2))

plot(visit_data$maritalstatus, visit_data$losdays2)
```
 not
 
```{r insurancetype}
insurancetype_lm = lm(losdays2~insurancetype)
summary(insurancetype_lm)
anova(insurancetype_lm)
par(mfrow=c(2,2))
plot(insurancetype_lm)
```

```{r bmi}
bmi_lm = lm(losdays2~bmi)
summary(bmi_lm)
par(mfrow=c(2,2))
plot(bmi_lm)
```


```{r mews}

mews_lm = lm(losdays2~mew_data)
summary(mews_lm)
par(mfrow=c(2,2))
plot(mews_lm)
patient_data1 %>%
  ggplot(aes(x = mews, y = heartrate)) + geom_point()
```
0-1=normal, 2-3=increase caution, 4-5=further deterioration, >5 immediate action required 

Below are different signs in MEWs
```{r bpsystolic}
library(stats)
bpsystolic_lm = lm(losdays2~bpsystolic)
summary(bpsystolic_lm)
par(mfrow=c(2,2))
plot(bpsystolic_lm)
plot(bpsystolic)
a = rstandard(bpsystolic_lm)
b = a[abs(a)>2.5]
```

```{r o2sat}
o2sat_lm = lm(losdays2~o2sat)
summary(o2sat_lm)
par(mfrow=c(2,2))
plot(o2sat_lm)  ## normal using percentage
```

```{r temperature}
temperature_lm = lm(losdays2~temperature)
summary(temperature_lm)
par(mfrow=c(2,2))
plot(temperature_lm)
plot(temperature) ## exceed 50
patient_data1 %>%
  ggplot(aes(x = patientid, y = evisit)) + geom_point()
```
not significant


```{r heartrate}
heartrate_lm = lm(losdays2~heartrate)
summary(heartrate_lm)
par(mfrow=c(2,2))
plot(heartrate_lm)
patient_data1 %>%
  ggplot(aes(x = patientid, y = heartrate)) + geom_point() ## has heart rate more than 200
```

```{r respirationrate}
respirationrate_lm = lm(losdays2~respirationrate)
summary(respirationrate_lm)
par(mfrow=c(2,2))
plot(respirationrate_lm)
hist(respirationrate)
```

```{r bpdiastolic}
bpdiastolic_lm = lm(losdays2~bpdiastolic)
summary(bpdiastolic_lm)
par(mfrow=c(2,2))
plot(bpdiastolic_lm)  
patient_data1 %>%
  ggplot(aes(x = patientid, y = bpdiastolic)) + geom_point()
```




**1.Forward elimination procedures**
```{r}
model <- lm(losdays2 ~ is30dayreadmit + mews + cindex + bmi + evisit + icu_flag + ageyear + gender + race + religion + maritalstatus + insurancetype, data = patient_data1)
ols_step_forward(model, details = TRUE)

# result: ageyear is30dayreadmit gender insurancetype race bmi
```


**2.Backward elimination procedures**
```{r}
model <- lm(losdays2 ~ is30dayreadmit + mews + cindex + bmi + evisit + icu_flag + ageyear + gender + race + religion + maritalstatus + insurancetype, data = patient_data1)
ols_best_subset(model)
```

**3.Stepwise regression procedures**
```{r}
model <- lm(losdays2 ~ is30dayreadmit + mews + cindex + bmi + evisit + icu_flag + ageyear + gender + race + religion + maritalstatus + insurancetype, data = patient_data1)
ols_stepwise(model, details = FALSE)
##not work
```

**AIC procedure
```{r}
ols_stepaic_forward(model)  #cindex mews is30dayreadmit gender insurancetype ageyear
ols_stepaic_backward(model) #is30dayreadmit mews cindex evisit ageyear  gender  insurancetype
ols_stepaic_both(model) #cindex evisit is30dayreadmit gender insurancetype ageyear

ols_step_backward(model, details = TRUE)
```





```{r}
# Saving dataset for automatic search procedures in SAS

#write_csv(patient_data1, "/Users/mhk/Desktop/patient_data1.csv")

# patient_data1 = patient_data1 %>%
#   mutate(male = ifelse(gender == "Male", 1, 0),
#          afr_american = ifelse(race == "African Amer/Black", 1, 0),
#          asian = ifelse(race == "Asian", 1, 0),
#          nat_american = ifelse(race == "Native Amer/Alaskan", 1, 0),
#          nat_hawaii = ifelse(race == "Natv Hawaii/Pacf Isl", 1, 0),
#          multi_racial = ifelse(race == "Other/Multiracial", 1, 0),
#          angelican = ifelse(religion == "Angelican", 1, 0),
#          catholic = ifelse(religion == "Catholic", 1, 0),
#          christian = ifelse(religion == "Christian", 1, 0),
#          hebrew = ifelse(religion == "Hebrew", 1, 0),
#          hindu = ifelse(religion == "Hindu", 1, 0),
#          islam = ifelse(religion == "Islam", 1, 0),
#          jewish = ifelse(religion == "Jewish", 1, 0),
#          mormon = ifelse(religion == "Mormon", 1, 0),
#          non_denom = ifelse(religion == "Non Denominational", 1, 0),
#          married = ifelse(religion == "Married", 1, 0),
#          widowed = ifelse(maritalstatus == "Widowed", 1, 0),
#          divorced = ifelse(religion == "Divorced", 1, 0),
#          separated = ifelse(religion == "Separated", 1, 0),
#          
#          )
  
  # mutate(gender = recode(gender, "Female" = 1, "Male" = 0),
  #        race = recode(race, "White" = 0, "African Amer/Black" = 1, "Asian" = 2, "Native Amer/Alaskan" = 3, "Natv Hawaii/Pacf Isl" = 4, "Other/Multiracial" = 5),
  #        religion = recode(religion, "No Affiliation" = 0, "Angelican" = 1,  "Catholic" = 2, "Christian" = 3, "Hebrew" = 4, "Hindu" = 5, "Islam" = 6, "Jewish" = 7, "Mormon" = 8, "Non Denominational" = 9, "Other" = 10),
  #        maritalstatus = recode(maritalstatus, "Single" = 0, "Married" = 1, "Widowed" = 2, "Divorced" = 3, "Separated" = 4, "Civil Union" = 5),
  #       insurancetype = recode(insurancetype, "Private" = 0, "Medicare" = 1, "Medicaid" = 2)) %>%
  # write_csv(., "/Users/mhk/Desktop/patient_data1.csv")

```


```{r}
library("leaps")
model1 = lm(losdays2 ~  mews + cindex + evisit  + ageyear  + insurancetype, data = patient_data1)
model2 = lm(losdays2 ~ is30dayreadmit + mews + cindex + ageyear + gender +  insurancetype, data = patient_data1)
model3 = lm(losdays2 ~ is30dayreadmit + cindex +  evisit +  ageyear + gender +  insurancetype, data = patient_data1)
model4 = lm(losdays2 ~ mews + cindex + bmi + evisit + ageyear + insurancetype, data = patient_data1)
model5 = lm(losdays2 ~ is30dayreadmit +  bmi +  ageyear + gender + race + insurancetype, data = patient_data1)
model6 = lm(losdays2 ~ is30dayreadmit + mews + cindex +  evisit + ageyear + gender  + insurancetype, data = patient_data1)
model7 = lm(losdays2 ~ mews + cindex + bmi + evisit  + ageyear + maritalstatus + insurancetype, data = patient_data1)
model8 = lm(losdays2 ~ mews + cindex + bmi + evisit +  ageyear + gender  + religion + maritalstatus + insurancetype, data = patient_data1)
anova(model4, model7)
AIC(model5)
AIC(model6)
AIC(model8)

```

```{r}
par(mfrow=c(2,2))
plot(model8)
```





